Bayesian Rate Estimation

### 2012-08-13 many.tex
Spotted two more typos.

Putting everything together, the posterior is
\begin{equation}
  \label{eq:posterior}
  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto \left[
    \prod_{\mathset{i | f_i = 1}} R_f \hat{f}\left(x_i, \theta\right) \right]
  \left[ \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
  \right] \exp\left[ - \left(R_f + R_b\right) \right]
  \frac{p(\theta)}{\sqrt{R_f R_b}}
\end{equation}
Not surprisingly, this posterior is equivalent to one derived by
assuming that the flags, $\mathset{f_i}$, are un-observed data and
treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
  0}$ as samples from an inhomogeneous Poisson process.  For an
inhomogeneous Poisson process with rate function $r(y)$ (cumulative
rate $R(y)$), the likelihood of a set of samples $\mathset{y_i}$ is
given by 
\begin{multline}
  p\left( \mathset{y_i} | r \right) = \lim_{\epsilon \to 0} P\left(
  \textnormal{zero events below } y_0 - \epsilon \right) \times
  P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{
    and } y_0 + \epsilon \right) \\ \times P\left( \textnormal{zero
    events between } y_0+\epsilon \textnormal{ and } y_1 - \epsilon
  \right) \ldots \\ = \lim_{\epsilon \to 0} \exp\left[ - R\left(y_0 -
    \epsilon \right) \right] \left[ r\left( y_0 \right) +
    \order{\epsilon^2} \right] \times \exp\left[ - \left[ R\left( y_1
      - \epsilon\right) - R\left( y_0 + \epsilon\right) \right]
    \right] \times \ldots = \left[\prod_i r\left( y_i
    \right)\right] \exp\left[ - R\left( \infty \right) \right].
\end{multline}
Applying this likelihood once to the foreground samples and once to
the background samples reproduces Eq.~\eqref{eq:posterior}.

@@

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{2}
      \right)}{2} \right)^N
\end{equation}
Imposing the threshold, $x > \xmin$, the
cumulative distribution of the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
    \right)^N }
\end{equation}
\ilya{for $x>\xmin$, $0$ otherwise.}

@@

To analyize our synthetic data set, we analytically marginalized over
the state flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
\ilya{[I think it would be kind to the reader to be more explicit
    about the connection with the formalism in Section II.  What are
    the "events"?  What is the ranking statistic $x$?  What are the
    flags (which aren't in $\mathbb{R}$?)?]}  We did this to take
advantage of the \textit{emcee} sampler of \citet{ForemanMackey2012},
which requires all parameters to be in $\mathbb{R}$.  We applied a
prior on the shape parameters that is flat in $\vec{x}_0$ and
$\vec{\gamma}$, and the Jeffreys prior on $r_0$,
\begin{equation}
  p\left( r_0 \right) = \frac{\sqrt{R_f}}{r_0}.
\end{equation}
\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
    why the prior on $r_0$ should depend on $R_f$?]}  (Note that this
factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
$1/\sqrt{R_f}$; we have verified that the priors on these parameters
are irrelevant to our results, as would be expected from the
measurement of $\sim 1000$ foreground stars.)

>>>>

Putting everything together, the posterior is
\begin{equation}
  \label{eq:posterior}
  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto \left[
    \prod_{\mathset{i | f_i = 1}} R_f \hat{f}\left(x_i, \theta\right) \right]
  \left[ \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
  \right] \exp\left[ - \left(R_f + R_b\right) \right]
  \frac{p(\theta)}{\sqrt{R_f R_b}}
\end{equation}
Not surprisingly, this posterior is equivalent to one derived by
assuming that the flags, $\mathset{f_i}$, are un-observed data and
treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
  0}$ as samples from an inhomogeneous Poisson process.  For an
inhomogeneous Poisson process with rate function $r(y)$ (cumulative
rate $R(y)$), the likelihood of a set of samples $\mathset{y_i}$ is
given by 
\begin{multline}
  p\left( \mathset{y_i} | r \right) = \lim_{\epsilon \to 0} P\left(
  \textnormal{zero events below } y_0 - \epsilon \right) \times
  P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{
    and } y_0 + \epsilon \right) \\ \times P\left( \textnormal{zero
    events between } y_0+\epsilon \textnormal{ and } y_1 - \epsilon
  \right) \ldots \\ = \lim_{\epsilon \to 0} \exp\left[ - R\left(y_0 -
    \epsilon \right) \right] \left[ r\left( y_0 \right) +
    \order{\epsilon^2} \right] \times \exp\left[ - \left[ R\left( y_1
      - \epsilon\right) - R\left( y_0 + \epsilon\right) \right]
    \right] \times \ldots = \left[\prod_i r\left( y_i
    \right)\right] \exp\left[ - R\left( \infty \right) \right].
\end{multline}
\ilya{[Note: I don't think this is right -- the expression you've written down should be zero in the limit $\epsilon \to 0$, and it would be if the $\epsilon$s were kept correctly (e.g., $ P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{ and } y_0 + \epsilon \right) \sim  r\left( y_0 \right) \epsilon +
    \order{\epsilon^2}$).]}
Applying this likelihood once to the foreground samples and once to
the background samples reproduces Eq.~\eqref{eq:posterior}.

@@

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
      \right)}{2} \right)^N
\end{equation}
Imposing the threshold, $x > \xmin$, the
cumulative distribution of the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}} \right)
    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
    \right)^N }
\end{equation}
\ilya{for $x>\xmin$, $0$ otherwise.}

@@

To analyze our synthetic data set, we analytically marginalized over
the state flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
\ilya{[I think it would be kind to the reader to be more explicit
    about the connection with the formalism in Section II.  What are
    the "events"?  What is the ranking statistic $x$?  What are the
    flags (which aren't in $\mathbb{R}$?)?]}  We did this to take
advantage of the \textit{emcee} sampler of \citet{ForemanMackey2012},
which requires all parameters to be in $\mathbb{R}$.  We applied a
prior on the shape parameters that is flat in $\vec{x}_0$ and
$\vec{\gamma}$, and the Jeffreys prior on $r_0$,
\begin{equation}
  p\left( r_0 \right) = \frac{\sqrt{R_f}}{r_0}.
\end{equation}
\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
    why the prior on $r_0$ should depend on $R_f$?]}  (Note that this
factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
$1/\sqrt{R_f}$; we have verified that the priors on these parameters
are irrelevant to our results, as would be expected from the
measurement of $\sim 1000$ foreground stars.)
